---
title: "Merge"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(opalr)
#opal_url <- "http://localhost:8080"
opal_url <- "https://opal-demo.obiba.org"
```

## Method 1

In this method, all data processing is done in the R client session.

### Extract Datasets

Retrieve the data stored in Opal as tibbles in the local R session.

```{r extract}
o <- opal.login(username = "administrator", password = "password", url = opal_url)
lab <- harmo.table_get(o, "lumc", "lab")
medication <- harmo.table_get(o, "lumc", "medication")
patient <- harmo.table_get(o, "lumc", "patient")
```

### Merge Datasets

Merge by joining datasets. See tutorial on how to [Join in R](http://www.datasciencemadesimple.com/join-in-r-merge-in-r/).

You need to choose what kind of join makes more sense for you:

* inner join:

```{r inner_med}
inner_medication <- merge(x = medication, y = patient, by = "id")
inner_medication
```

* outer join:

```{r outer_med}
outer_medication <- merge(x = medication, y = patient, by = "id", all = TRUE)
outer_medication
```

For the lab results, you need to define same **id** column in both datasets that will represent the patient ID:

```{r prepare_lab}
# rename id
names(lab)[1] <- "sample_id"
# rename patient_id
names(lab)[2] <- "id"
lab
```

Then do the join:

* inner join:

```{r inner_lab}
inner_lab <- merge(x = lab, y = patient, by = "id")
inner_lab
```

* outer join:

```{r outer_lab}
outer_lab <- merge(x = lab, y = patient, by = "id", all = TRUE)
outer_lab
```


### Save Datasets

Prepare datasets dictionaries.

```{r dico}
# make tibbles instead of data.frames before saving
inner_medication <- tibble::as_tibble(inner_medication)
outer_medication <- tibble::as_tibble(outer_medication)
inner_lab <- tibble::as_tibble(inner_lab)
outer_lab <- tibble::as_tibble(outer_lab)

# note that it is possible to tweak a bit the data dictionary to help Opal in interpreting the data  
attributes(outer_lab$IL6) <- list("opal.occurrence_group" = "lab")
attributes(outer_lab$IL6_date) <- list("opal.occurrence_group" = "lab")
attributes(outer_lab$sample_id) <- list("opal.occurrence_group" = "lab")
attributes(outer_lab$Age) <- list("opal.repeatable" = 0)
```

Upload datasets and save.

```{r save}
harmo.table_save(o, inner_medication, "lumc", "inner_medication", overwrite = TRUE, force = TRUE)
harmo.table_save(o, outer_medication, "lumc", "outer_medication", overwrite = TRUE, force = TRUE)
harmo.table_save(o, inner_lab, "lumc", "inner_lab", overwrite = TRUE, force = TRUE)
harmo.table_save(o, outer_lab, "lumc", "outer_lab", overwrite = TRUE, force = TRUE)

# clean up
opal.logout(o)
```

## Method 2

In this method, all data processing is done in the R server session. It is less convenient when exploring data, but is preferable when systematically applying a data proccess because it avoids transferring data to/from client and server R sessions.

### Assign Datasets

Retrieve the data stored in Opal as tibbles in the local R session.

```{r assign}
o <- opal.login(username = "administrator", password = "password", url = opal_url)
opal.assign.table.tibble(o, "lab", "lumc.lab")
opal.assign.table.tibble(o, "medication", "lumc.medication")
opal.assign.table.tibble(o, "patient", "lumc.patient")
opal.symbols(o)
```
### Merge Datasets

Do merge operations:

```{r assign_merge}
# merge medication with patient
opal.assign.script(o, "patient_medication", quote(merge(x = medication, y = patient, by = "id", all = TRUE)))
# rename lab columns and merge lab with patient
opal.execute(o, 'names(lab)[1] <- "sample_id"')
opal.execute(o, 'names(lab)[2] <- "id"')
opal.assign.script(o, "patient_lab", quote(merge(x = lab, y = patient, by = "id", all = TRUE)))
```

Another option is to use the `opal.execute.source()` function:

```{r eval=FALSE}
# the merge.R file, located in the Opal file system contains all the data processing code
opal.execute.source(o, "/projects/lumc/merge.R")
```

### Save datasets

Upload datasets and save.

```{r save2}
opal.assign.script(o, "patient_medication", quote(tibble::as_tibble(patient_medication)))
opal.symbol_import(o, "patient_medication", "lumc")
opal.assign.script(o, "patient_lab", quote(tibble::as_tibble(patient_lab)))
opal.symbol_import(o, "patient_lab", "lumc")

# clean up
opal.logout(o)
```
